\thetitle\thanks{This work was published, in part, in Runtime Verification (RV) 2016~\cite{Shi2016}.}

### 2017-07-09 submitted/all-in-one.tex
comments

We present a new technique than enables validation of protocol implementations using
wireless sniffers. Given a monitor state machine representing the protocol being
validated, we describe a systematic transformation that adds non-deterministic
transitions to incorporate uncertainty introduced by the sniffer. This
augmented validation state machine implicitly defines a set of
mutated traces, each satisfying the original state machine with a specific
likelihood. If the set is empty, the implementation definitely violates the
protocol. Searching over all the mutated traces is NP-complete, but the approach
can be made practical by applying protocol-oblivious heuristics that limit the
search to likely mutated traces.

@@

A tuple $(t_i, p_i)$ in the packet trace means the packet $p_i$ is presented to
the state machine at time $t_i$. The monitor {\it rejects a trace} $Tr$ if there
exists a prefix of $Tr$ such that all states reachable after consuming the
prefix have no valid transitions for the next $(t, p)$ input.

As an example, the monitor state machine illustrated in
Fig.~\ref{fig:dot11_tx_ta} can be formally defined as follows:

\begin{itemize}
  \item $\Sigma = \{\it{DATA}_i, \it{DATA}'_i, Ack\ |\ 0 \le i < N\}$.
  \item Clock variables $C = \{c\}$. The only clock variable $c$ is
    used for acknowledgment time out.
  \item $\mathbb{X} = \{i\}$, as a variable with ${\mathit log}(N) + 1$ bits to
    count from $0$ to $N$.  
  \item Guard constraints $G = \{ c \le T_o, c > T_o, T_o < c \le T_m\}$.
    $T_o$ is the acknowledgment time out value, and $T_m >
    T_o$ is the maximum delay allowed before the retransmission packet gets
    sent. $T_o$ can be arbitrary large but not infinity in order to check the
    liveness of the DUT.
\end{itemize}

@@

@@

\begin{comment}
\textbf{Sniffer Trace Analysis.} Wireless sniffers has been widely used to
analyze MAC layer behaviors of enterprise wireless
networks~\cite{sheng:wicom2008,tan:tmc2014,yeo-wise04,yeo:witmemo2005}.
Jigsaw~\cite{Cheng:2006:JSP:1159913.1159920} is a larger scale wireless network
monitoring infrastructure. 150~radio monitors were deployed in a campus
building. Traces collected from multiple sniffers were merged and synchronized
to restructure the link and transportation layer conversations. Protocol
specific heuristics were developed to infer the missing packets. The work
in~\cite{Mahajan:2006:AMB:1159913.1159923} shared the same idea of trace merging
with Jigsaw, but uses a FSM to infer packet reception. These works assume the
correctness of the protocol implementation in order to infer missing packets,
while we systematically encode the uncertainty of sniffer traces for
verification purpose.

\textbf{ Testing Under Uncertainty}. The position paper by Rosemblum \textit{et
al}~\cite{Elbaum:2014:KUT:2635868.2666608} contains excellent motivation for the
need to combat uncertainty foundationally when testing systems. McKinley
\textit{et al}~\cite{bornholt2014uncertain,sampson2014expressing} deals with checking
assertions in programs dealing with noisy data from sensors. Instead of checking
the truth or falsity of assertions, they model the probability distribution of
the assertion conditions and perform Monte-carlo based simulations to estimate
the probabilities. Our work can be seen as leveraging non-determinism to weaken
the specification logically to precisely define the problem complexity, and use
probabilities to guide the search for likely mutations.  Other works have used
sampling to find data-race bugs~\cite{marino2009literace}, and ensure that the
sampling does not lead to spurious alarms.
\end{comment}

@@

We formally define the uncertainty problem in validating wireless protocol
implementations using sniffers. We describe a systematic augmentation of the
protocol state machine to explicitly encode the uncertainty of sniffer traces.
We characterize the NP-completeness of the problem and propose both an
exhaustive search algorithm and heuristics to restrict the search to more likely
traces. We present two case studies using \ns{} network simulator to demonstrate
how our framework can improve validation precision and detect real bugs. We also
report the application of our framework on a commercial product under
development.

@@

@@

@@

@@

>>>>

We present a new technique than enables validation of protocol implementations using
wireless sniffers. Given a monitor state machine representing the protocol being
validated, we describe a systematic transformation that adds non-deterministic
transitions to incorporate uncertainty introduced by the sniffer. This
augmented validation state machine implicitly defines a set of
mutated traces, each satisfying the original state machine with a specific
likelihood. If the set of mutated traces is empty, the implementation definitely violates the
protocol. Searching over all the mutated traces is NP-complete, but the approach
can be made practical by applying protocol-oblivious heuristics that limit the
search to likely mutated traces.

@@

A tuple $(t_i, p_i)$ in the packet trace means the packet $p_i$ is presented to
the state machine at time $t_i$.  As an example, the monitor state machine
illustrated in Fig.~\ref{fig:dot11_tx_ta} can be formally defined as follows:

\begin{itemize}
  \item $\Sigma = \{\it{DATA}_i, \it{DATA}'_i, Ack\ |\ 0 \le i < N\}$.
  \item Clock variables $C = \{c\}$. The only clock variable $c$ is
    used for acknowledgment time out.
  \item $\mathbb{X} = \{i\}$, as a variable with ${\mathit log}(N) + 1$ bits to
    count from $0$ to $N$.  
  \item Guard constraints $G = \{ c \le T_o, c > T_o, T_o < c \le T_m\}$.
    $T_o$ is the acknowledgment time out value, and $T_m >
    T_o$ is the maximum delay allowed before the retransmission packet gets
    sent. In this particular case, $T_o$ can be arbitrary large but not infinity
        in order to check the liveness of the DUT.
\end{itemize}

\begin{definition}[Trace Rejection and Acceptance]
    A monitor state machine $S$ {\it rejects a trace} $Tr$ if there
    exists a prefix of $Tr$ such that all states reachable after consuming the
    prefix have no valid transitions for the next $(t, p)$ input. Otherwise, $S$
    {\it accepts} $Tr$ (or $Tr$ {\it satisfies} $S$).
\end{definition}

@@

The DUT and endpoint are configured to use the IEEE 802.11g standard with both
RTS/CTS and fragmentation disabled.  A Constant Bit Rate (CBR) UDP traffic
(54~Mbps) is generated from the DUT to the endpoint.  The UDP packet size is
1436 bytes, which results in a 1500 bytes \wifi{} packet.

In order to control the packet loss ratios between each pair of devices, we
developed a new propagation loss model for \ns{} called
\texttt{MatrixRandomPropagationLossModel}.
\sloppy{  Instead of a constant propagation
  loss as in existing \texttt{MatrixPropagationLossModel}, the signal
  propagation loss between a pair of nodes is determined by a binary random
  variable of two  values: 0~dB (no loss) and 1000~dB (complete loss).
  }
Therefore, any packet loss
probability can be achieved by adjusting the random variable distribution.
The
model supports both symmetric and asymmetric propagation losses.
We use
symmetric propagation loss in all our experiments.
Finally, pcap capture is enabled in both the DUT and the sniffer devices.

@@

\begin{comment}
\textbf{Sniffer Trace Analysis.} Wireless sniffers has been widely used to
analyze MAC layer behaviors of enterprise wireless
networks~\cite{sheng:wicom2008,tan:tmc2014,yeo-wise04,yeo:witmemo2005}.
Jigsaw~\cite{Cheng:2006:JSP:1159913.1159920} is a larger scale wireless network
monitoring infrastructure. 150~radio monitors were deployed in a campus
building. Traces collected from multiple sniffers were merged and synchronized
to restructure the link and transportation layer conversations. Protocol
specific heuristics were developed to infer the missing packets. The work
in~\cite{Mahajan:2006:AMB:1159913.1159923} shared the same idea of trace merging
with Jigsaw, but uses a FSM to infer packet reception. These works assume the
correctness of the protocol implementation in order to infer missing packets,
while we systematically encode the uncertainty of sniffer traces for
verification purpose.
\end{comment}

\textbf{ Testing Under Uncertainty}. The position paper by Rosemblum \textit{et
al}~\cite{Elbaum:2014:KUT:2635868.2666608} contains excellent motivation for the
need to combat uncertainty foundationally when testing systems. McKinley
\textit{et al}~\cite{bornholt2014uncertain,sampson2014expressing} deals with checking
assertions in programs dealing with noisy data from sensors. Instead of checking
the truth or falsity of assertions, they model the probability distribution of
the assertion conditions and perform Monte-carlo based simulations to estimate
the probabilities. Our work can be seen as leveraging non-determinism to weaken
the specification logically to precisely define the problem complexity, and use
probabilities to guide the search for likely mutations.  Other works have used
sampling to find data-race bugs~\cite{marino2009literace}, and ensure that the
sampling does not lead to spurious alarms.

@@

We formally define an instance of the uncertainty problem in validating wireless protocol
implementations using sniffers. We describe a systematic augmentation of the
protocol state machine to explicitly encode the uncertainty of sniffer traces.
We characterize the NP-completeness of the problem and propose both an
exhaustive search algorithm and heuristics to restrict the search to more likely
traces. We present two case studies using \ns{} network simulator to demonstrate
how our framework can improve validation precision and detect real bugs. We also
report the application of our framework on a commercial product under
development.

@@

\bibitem{bornholt2014uncertain}
J.~Bornholt, T.~Mytkowicz, and K.~S. McKinley.
\newblock Uncertain< t>: A first-order type for uncertain data.
\newblock {\em ACM SIGARCH Computer Architecture News}, 42(1):51--66, 2014.

@@

\bibitem{Elbaum:2014:KUT:2635868.2666608}
S.~Elbaum and D.~S. Rosenblum.
\newblock Known unknowns: Testing in the presence of uncertainty.
\newblock In {\em Proceedings of the 22Nd ACM SIGSOFT International Symposium
  on Foundations of Software Engineering}, FSE 2014, pages 833--836, New York,
  NY, USA, 2014. ACM.

@@

\bibitem{marino2009literace}
D.~Marino, M.~Musuvathi, and S.~Narayanasamy.
\newblock Literace: effective sampling for lightweight data-race detection.
\newblock In {\em ACM Sigplan Notices}, volume~44, pages 134--143. ACM, 2009.

@@

\bibitem{sampson2014expressing}
A.~Sampson, P.~Panchekha, T.~Mytkowicz, K.~S. McKinley, D.~Grossman, and
  L.~Ceze.
\newblock Expressing and verifying probabilistic assertions.
\newblock In {\em ACM SIGPLAN Notices}, volume~49, pages 112--122. ACM, 2014.
